### 2020-10-08 analysis/analysis.Rmd
starting to clean up data

---
title: "Dragonfly Territory Ananlysis"
Author: Maggie Vincent
Term: Fall 2020
Course: Reproducible Research in R
output: pdf_document
---

# Read in Data from BioTime

rawDragonfly <- `rawdata.(1)`

install.packages('plyr')
install.packages('dplyr')

library('plyr')
library('dplyr')

coenigrion <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Coenagrion puella')
ischnura <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Ischnura elegans')
lestes <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Lestes sponsa')
libelsympetrum <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Libellula quadrimaculata')
sympetrum <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Sympetrum striolatum')

### 2020-10-29 analysis/analysis.Rmd
subsetting data.frames

@@
 title: "Dragonfly Territory Ananlysis"
-Author: Maggie Vincent
+Author: "Maggie Vincent"
 Term: Fall 2020
@@
 sympetrum <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Sympetrum striolatum')
+
+# Make a rough plot to visualize the data
+
+par(mfrow=c(2,3))
+plot(ischnura$YEAR, ischnura$ABUNDANCE)
+plot(coenigrion$YEAR, coenigrion$ABUNDANCE)
+plot(lestes$YEAR, lestes$ABUNDANCE)
+plot(libellula$YEAR, libellula$ABUNDANCE)
+plot(sympetrum$YEAR, sympetrum$ABUNDANCE)
+
+# I don't like that the y-axes are scaled differently and I want to change the axis labels
+
+# subsetting down the data.frames, since I need to do it more than once, I'll write a simple function
+
+getAbundanceYear <- function(dat, x = 'ABUNDANCE', y = 'YEAR'){
+                    tmp <- subset(dat, select = c(x, y))
+                    return(tmp)
+}
+
+ischnuraAY <- getAbundanceYear(ischnura)
+coenigrionAY <- getAbundanceYear(coenigrion)
+lestesAY <- getAbundanceYear(lestes)
+libellulaAY <- getAbundanceYear(libellula)
+sympetrumAY <- getAbundanceYear(sympetrum)
+
+# the current data.frames are different lengths
+# this is because the years without observations of the species were not put in as zeroes
+# after consultation with Tad Dallas, I will be changing the data.frames to length 27 (observation years) with zeroes for the missing years

### 2020-11-13 LICENSE.Rmd
had to reclone for commits to go through from local

---
title: "LICENSE"
output: pdf_document
---

Creative Commons Legal Code

CCO Attribution 4.0 International

Creative Commons Corporation (“Creative Commons”) is not a law firm and does not provide legal services or legal advice. Distribution of Creative Commons public licenses does not create a lawyer-client or other relationship. Creative Commons makes its licenses and related information available on an “as-is” basis. Creative Commons gives no warranties regarding its licenses, any material licensed under their terms and conditions, or any related information. Creative Commons disclaims all liability for damages resulting from their use to the fullest extent possible.

By exercising the Licensed Rights (defined below), You accept and agree to be bound by the terms and conditions of this Creative Commons Attribution 4.0 International Public License ("Public License"). To the extent this Public License may be interpreted as a contract, You are granted the Licensed Rights in consideration of Your acceptance of these terms and conditions, and the Licensor grants You such rights in consideration of benefits the Licensor receives from making the Licensed Material available under these terms and conditions.

Section 1 – Definitions.

Adapted Material means material subject to Copyright and Similar Rights that is derived from or based upon the Licensed Material and in which the Licensed Material is translated, altered, arranged, transformed, or otherwise modified in a manner requiring permission under the Copyright and Similar Rights held by the Licensor. For purposes of this Public License, where the Licensed Material is a musical work, performance, or sound recording, Adapted Material is always produced where the Licensed Material is synched in timed relation with a moving image.
Adapter's License means the license You apply to Your Copyright and Similar Rights in Your contributions to Adapted Material in accordance with the terms and conditions of this Public License.
Copyright and Similar Rights means copyright and/or similar rights closely related to copyright including, without limitation, performance, broadcast, sound recording, and Sui Generis Database Rights, without regard to how the rights are labeled or categorized. For purposes of this Public License, the rights specified in Section 2(b)(1)-(2) are not Copyright and Similar Rights.
Effective Technological Measures means those measures that, in the absence of proper authority, may not be circumvented under laws fulfilling obligations under Article 11 of the WIPO Copyright Treaty adopted on December 20, 1996, and/or similar international agreements.
Exceptions and Limitations means fair use, fair dealing, and/or any other exception or limitation to Copyright and Similar Rights that applies to Your use of the Licensed Material.
Licensed Material means the artistic or literary work, database, or other material to which the Licensor applied this Public License.
Licensed Rights means the rights granted to You subject to the terms and conditions of this Public License, which are limited to all Copyright and Similar Rights that apply to Your use of the Licensed Material and that the Licensor has authority to license.
Licensor means the individual(s) or entity(ies) granting rights under this Public License.
Share means to provide material to the public by any means or process that requires permission under the Licensed Rights, such as reproduction, public display, public performance, distribution, dissemination, communication, or importation, and to make material available to the public including in ways that members of the public may access the material from a place and at a time individually chosen by them.
Sui Generis Database Rights means rights other than copyright resulting from Directive 96/9/EC of the European Parliament and of the Council of 11 March 1996 on the legal protection of databases, as amended and/or succeeded, as well as other essentially equivalent rights anywhere in the world.
You means the individual or entity exercising the Licensed Rights under this Public License. Your has a corresponding meaning.
Section 2 – Scope.

License grant.
Subject to the terms and conditions of this Public License, the Licensor hereby grants You a worldwide, royalty-free, non-sublicensable, non-exclusive, irrevocable license to exercise the Licensed Rights in the Licensed Material to:
reproduce and Share the Licensed Material, in whole or in part; and
produce, reproduce, and Share Adapted Material.
Exceptions and Limitations. For the avoidance of doubt, where Exceptions and Limitations apply to Your use, this Public License does not apply, and You do not need to comply with its terms and conditions.
Term. The term of this Public License is specified in Section 6(a).
Media and formats; technical modifications allowed. The Licensor authorizes You to exercise the Licensed Rights in all media and formats whether now known or hereafter created, and to make technical modifications necessary to do so. The Licensor waives and/or agrees not to assert any right or authority to forbid You from making technical modifications necessary to exercise the Licensed Rights, including technical modifications necessary to circumvent Effective Technological Measures. For purposes of this Public License, simply making modifications authorized by this Section 2(a)(4) never produces Adapted Material.
Downstream recipients.
Offer from the Licensor – Licensed Material. Every recipient of the Licensed Material automatically receives an offer from the Licensor to exercise the Licensed Rights under the terms and conditions of this Public License.
No downstream restrictions. You may not offer or impose any additional or different terms or conditions on, or apply any Effective Technological Measures to, the Licensed Material if doing so restricts exercise of the Licensed Rights by any recipient of the Licensed Material.
No endorsement. Nothing in this Public License constitutes or may be construed as permission to assert or imply that You are, or that Your use of the Licensed Material is, connected with, or sponsored, endorsed, or granted official status by, the Licensor or others designated to receive attribution as provided in Section 3(a)(1)(A)(i).
Other rights.

Moral rights, such as the right of integrity, are not licensed under this Public License, nor are publicity, privacy, and/or other similar personality rights; however, to the extent possible, the Licensor waives and/or agrees not to assert any such rights held by the Licensor to the limited extent necessary to allow You to exercise the Licensed Rights, but not otherwise.
Patent and trademark rights are not licensed under this Public License.
To the extent possible, the Licensor waives any right to collect royalties from You for the exercise of the Licensed Rights, whether directly or through a collecting society under any voluntary or waivable statutory or compulsory licensing scheme. In all other cases the Licensor expressly reserves any right to collect such royalties.
Section 3 – License Conditions.

Your exercise of the Licensed Rights is expressly made subject to the following conditions.

Attribution.

If You Share the Licensed Material (including in modified form), You must:

retain the following if it is supplied by the Licensor with the Licensed Material:
identification of the creator(s) of the Licensed Material and any others designated to receive attribution, in any reasonable manner requested by the Licensor (including by pseudonym if designated);
a copyright notice;
a notice that refers to this Public License;
a notice that refers to the disclaimer of warranties;
a URI or hyperlink to the Licensed Material to the extent reasonably practicable;
indicate if You modified the Licensed Material and retain an indication of any previous modifications; and
indicate the Licensed Material is licensed under this Public License, and include the text of, or the URI or hyperlink to, this Public License.
You may satisfy the conditions in Section 3(a)(1) in any reasonable manner based on the medium, means, and context in which You Share the Licensed Material. For example, it may be reasonable to satisfy the conditions by providing a URI or hyperlink to a resource that includes the required information.
If requested by the Licensor, You must remove any of the information required by Section 3(a)(1)(A) to the extent reasonably practicable.
If You Share Adapted Material You produce, the Adapter's License You apply must not prevent recipients of the Adapted Material from complying with this Public License.
Section 4 – Sui Generis Database Rights.

Where the Licensed Rights include Sui Generis Database Rights that apply to Your use of the Licensed Material:

for the avoidance of doubt, Section 2(a)(1) grants You the right to extract, reuse, reproduce, and Share all or a substantial portion of the contents of the database;
if You include all or a substantial portion of the database contents in a database in which You have Sui Generis Database Rights, then the database in which You have Sui Generis Database Rights (but not its individual contents) is Adapted Material; and
You must comply with the conditions in Section 3(a) if You Share all or a substantial portion of the contents of the database.
For the avoidance of doubt, this Section 4 supplements and does not replace Your obligations under this Public License where the Licensed Rights include other Copyright and Similar Rights.
Section 5 – Disclaimer of Warranties and Limitation of Liability.

Unless otherwise separately undertaken by the Licensor, to the extent possible, the Licensor offers the Licensed Material as-is and as-available, and makes no representations or warranties of any kind concerning the Licensed Material, whether express, implied, statutory, or other. This includes, without limitation, warranties of title, merchantability, fitness for a particular purpose, non-infringement, absence of latent or other defects, accuracy, or the presence or absence of errors, whether or not known or discoverable. Where disclaimers of warranties are not allowed in full or in part, this disclaimer may not apply to You.
To the extent possible, in no event will the Licensor be liable to You on any legal theory (including, without limitation, negligence) or otherwise for any direct, special, indirect, incidental, consequential, punitive, exemplary, or other losses, costs, expenses, or damages arising out of this Public License or use of the Licensed Material, even if the Licensor has been advised of the possibility of such losses, costs, expenses, or damages. Where a limitation of liability is not allowed in full or in part, this limitation may not apply to You.
The disclaimer of warranties and limitation of liability provided above shall be interpreted in a manner that, to the extent possible, most closely approximates an absolute disclaimer and waiver of all liability.
Section 6 – Term and Termination.

This Public License applies for the term of the Copyright and Similar Rights licensed here. However, if You fail to comply with this Public License, then Your rights under this Public License terminate automatically.
Where Your right to use the Licensed Material has terminated under Section 6(a), it reinstates:

automatically as of the date the violation is cured, provided it is cured within 30 days of Your discovery of the violation; or
upon express reinstatement by the Licensor.
For the avoidance of doubt, this Section 6(b) does not affect any right the Licensor may have to seek remedies for Your violations of this Public License.
For the avoidance of doubt, the Licensor may also offer the Licensed Material under separate terms or conditions or stop distributing the Licensed Material at any time; however, doing so will not terminate this Public License.
Sections 1, 5, 6, 7, and 8 survive termination of this Public License.
Section 7 – Other Terms and Conditions.

The Licensor shall not be bound by any additional or different terms or conditions communicated by You unless expressly agreed.
Any arrangements, understandings, or agreements regarding the Licensed Material not stated herein are separate from and independent of the terms and conditions of this Public License.
Section 8 – Interpretation.

For the avoidance of doubt, this Public License does not, and shall not be interpreted to, reduce, limit, restrict, or impose conditions on any use of the Licensed Material that could lawfully be made without permission under this Public License.
To the extent possible, if any provision of this Public License is deemed unenforceable, it shall be automatically reformed to the minimum extent necessary to make it enforceable. If the provision cannot be reformed, it shall be severed from this Public License without affecting the enforceability of the remaining terms and conditions.
No term or condition of this Public License will be waived and no failure to comply consented to unless expressly agreed to by the Licensor.
Nothing in this Public License constitutes or may be interpreted as a limitation upon, or waiver of, any privileges and immunities that apply to the Licensor or You, including from the legal processes of any jurisdiction or authority.

### 2020-11-13 analysis/analysisSS.Rmd
had to reclone for commits to go through from local

---
title: ""
Author: "Maggie Vincent"
Term: Fall 2020
Course: Reproducible Research in R
output: pdf_document
---

# Read in Data from BioTime - Single study
## this was downloaded from 'http://biotime.st-andrews.ac.uk/taxa.php?10' on 4 October 2020
## study name == 'The development of dragonfly communities and the consequences of territorial behaviour'
## study ID == ''
## BioTIME does not have an API, making this necessary

rawDragonfly <- `DataSingleStudy`

install.packages('plyr')
install.packages('dplyr')

library('plyr')
library('dplyr')

coenigrion <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Coenagrion puella')
ischnura <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Ischnura elegans')
lestes <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Lestes sponsa')
libellula <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Libellula quadrimaculata')
sympetrum <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Sympetrum striolatum')

# Make a rough plot to visualize the data

par(mfrow=c(2,3))
plot(ischnura$YEAR, ischnura$ABUNDANCE)
plot(coenigrion$YEAR, coenigrion$ABUNDANCE)
plot(lestes$YEAR, lestes$ABUNDANCE)
plot(libellula$YEAR, libellula$ABUNDANCE)
plot(sympetrum$YEAR, sympetrum$ABUNDANCE)

# I don't like that the y-axes are scaled differently and I want to change the axis labels

# subsetting down the data.frames, since I need to do it more than once, I'll write a simple function

getAbundanceYear <- function(dat, x = 'ABUNDANCE', y = 'YEAR'){
                    tmp <- subset(dat, select = c(x, y))
                    return(tmp)
}

ischnuraAY <- getAbundanceYear(ischnura)
coenigrionAY <- getAbundanceYear(coenigrion)
lestesAY <- getAbundanceYear(lestes)
libellulaAY <- getAbundanceYear(libellula)
sympetrumAY <- getAbundanceYear(sympetrum)

# the current data.frames are different lengths
# this is because the years without observations of the species were not put in as zeroes
# after consultation with Tad Dallas, I will be changing the data.frames to length 27 (observation years) with zeroes for the missing years

row1959 <- c(0, 1959)
row1960 <- c(0, 1960)
row1961 <- c(0, 1961)
row1962 <- c(0, 1962)
row1963 <- c(0, 1963)

row1972 <- c(0, 1972)
row1974 <- c(0, 1974)

coenigronCompAY <- rbind(row1959, row1960, row1961, row1962, coenigrionAY)
ischnuraCompAY <- rbind(row1959, row1960, row1961, ischnuraAY)
lestesCompAY <- rbind(row1959, row1960, row1961, row1962, row1963, lestesAY)
# libellula will be dealt with differently because it is missing rows within the dataframe
sympetrumCompAY <- rbind(row1959, row1960, row1961, sympetrumAY)

prelibellulaCompAY <- rbind(row1960, row1972, row1974, libellulaAY)
# this will add the rows I want, but now in chronological order
libellulaCompAY <- dplyr::arrange(prelibellulaCompAY, YEAR)

# there are species missing from this, they need to be added separately

brachytronAbundance <- c(0,0,0,0,0,0,0,0,0,1,0,0,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
pyrrhosomaAbundance <- c(0,0,3,20,18,5,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
aeshnaAbundance <- c(0,0,0,0,1,2,3,1,1,2,3,1,5,1,2,0,0,0,1,1,1,0,2,1,1,3,1,2,2,2)

### 2020-11-14 analysis/analysisSS.Rmd
larval data re-added

@@
 
-# Read in Data from BioTime - Single study
+# Read in Data from BioTime - Adult Single study
+# An additional study with damselfly larval data will be read in the same way
 ## this was downloaded from 'http://biotime.st-andrews.ac.uk/taxa.php?10' on 4 October 2020
@@
 aeshnaAbundance <- c(0,0,0,0,1,2,3,1,1,2,3,1,5,1,2,0,0,0,1,1,1,0,2,1,1,3,1,2,2,2)
+
+# Damselfly Larval Data
+
+larval_data <- read.csv("~/GitHub/rrirfinalproject/larval_data.csv")

### 2020-11-23 analysis/analysisSS.Rmd
working on analysis

@@
 
-coenigrion <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Coenagrion puella')
-ischnura <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Ischnura elegans')
-lestes <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Lestes sponsa')
-libellula <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Libellula quadrimaculata')
-sympetrum <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Sympetrum striolatum')
+cPuella <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Coenagrion puella')
+iElegans <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Ischnura elegans')
+lSponsa <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Lestes sponsa')
+lQuadrimaculata <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Libellula quadrimaculata')
+sStriolatum <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Sympetrum striolatum')
 
@@
 par(mfrow=c(2,3))
-plot(ischnura$YEAR, ischnura$ABUNDANCE)
-plot(coenigrion$YEAR, coenigrion$ABUNDANCE)
-plot(lestes$YEAR, lestes$ABUNDANCE)
-plot(libellula$YEAR, libellula$ABUNDANCE)
-plot(sympetrum$YEAR, sympetrum$ABUNDANCE)
+plot(iElegans$YEAR, iElegans$ABUNDANCE)
+plot(cPuella$YEAR, cPuella$ABUNDANCE)
+plot(lSponsa$YEAR, lSponsa$ABUNDANCE)
+plot(lQuadrimaculata$YEAR, lQuadrimaculata$ABUNDANCE)
+plot(sStriolatum$YEAR, sStriolatum$ABUNDANCE)
 
@@
 
-ischnuraAY <- getAbundanceYear(ischnura)
-coenigrionAY <- getAbundanceYear(coenigrion)
-lestesAY <- getAbundanceYear(lestes)
-libellulaAY <- getAbundanceYear(libellula)
-sympetrumAY <- getAbundanceYear(sympetrum)
+iElegansAY <- getAbundanceYear(iElegans)
+cPuellaAY <- getAbundanceYear(cPuella)
+lSponsaAY <- getAbundanceYear(lSponsa)
+lQuadrimaculataAY <- getAbundanceYear(lQuadrimaculata)
+sStriolatumAY <- getAbundanceYear(sStriolatum)
+
+# I want to rename the abundance column with species name to aid comparison and readibility
+
+iEleganSAY <- dplyr::rename(iElegansAY, iElegans_abundance = ABUNDANCE)
+cPuellaSAY <- 
+lSponsaSAY <- 
+lQuadrimaculataSAY <- 
+sStriolatumSAY <-
 
@@
 
-coenigronCompAY <- rbind(row1959, row1960, row1961, row1962, coenigrionAY)
-ischnuraCompAY <- rbind(row1959, row1960, row1961, ischnuraAY)
-lestesCompAY <- rbind(row1959, row1960, row1961, row1962, row1963, lestesAY)
-# libellula will be dealt with differently because it is missing rows within the dataframe
-sympetrumCompAY <- rbind(row1959, row1960, row1961, sympetrumAY)
+cPuellaCompAY <- rbind(row1959, row1960, row1961, row1962, cPuellaSAY)
+iElegansCompAY <- rbind(row1959, row1960, row1961, iElegansSAY)
+lSponsaCompAY <- rbind(row1959, row1960, row1961, row1962, row1963, lSponsaSAY)
+# lQuadrimaculata will be dealt with differently because it is missing rows within the dataframe
+sStriolatumCompAY <- rbind(row1959, row1960, row1961, sStriolatumSAY)
 
-prelibellulaCompAY <- rbind(row1960, row1972, row1974, libellulaAY)
+prelibellulaCompAY <- rbind(row1960, row1972, row1974, lQuadrimaculataSAY)
 # this will add the rows I want, but now in chronological order
-libellulaCompAY <- dplyr::arrange(prelibellulaCompAY, YEAR)
+lQuadrimaculataCompAY <- dplyr::arrange(prelibellulaCompAY, YEAR)
 
-# there are species missing from this, they need to be added separately
+# there are species missing from this dataset, they need to be added separately, I will source (and cite) the abundance data
+# also need a vector corresponding to the years
 
-brachytronAbundance <- c(0,0,0,0,0,0,0,0,0,1,0,0,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
-pyrrhosomaAbundance <- c(0,0,3,20,18,5,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
-aeshnaAbundance <- c(0,0,0,0,1,2,3,1,1,2,3,1,5,1,2,0,0,0,1,1,1,0,2,1,1,3,1,2,2,2)
+dragonflyYears <- c(1959:1988)
+
+bPratenseAbundance <- c(0,0,0,0,0,0,0,0,0,1,0,0,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
+pNymphulaAbundance <- c(0,0,3,20,18,5,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
+aCyaneaAbundance <- c(0,0,0,0,1,2,3,1,1,2,3,1,5,1,2,0,0,0,1,1,1,0,2,1,1,3,1,2,2,2)
+sSanguineumAbundance <- c(0,0,0,0,0,1,0,0,1,0,2,0,1,0,0,1,0,1,4, 11, 3, 0, 1, 9, 6, 10, 4, 9, 3, 4)
+aGrandisAbundance <- c(0,0,0,0,0,0,0,4,0,1,1,0,2,0,2,0,0,0,1,1,0,1,2,1,0,1,1,4,0,1)
+
+bPratenseCompAY <- data.frame(bPratenseAbundance, dragonflyYears)
+pNymphulaCompAY <- data.frame(pNymphulaAbundance, dragonflyYears)
+aCyaneaCompAY <- data.frame(aCyaneaAbundance, dragonflyYears)
+sSanguineumCompAY <- data.frame(sSanguineumAbundance, dragonflyYears)
+aGrandisCompAY <- data.frame(aGrandisAbundance, dragonflyYears)
 

### 2020-11-24 analysis/analysisSS.Rmd
data manip work

@@
 
-iEleganSAY <- dplyr::rename(iElegansAY, iElegans_abundance = ABUNDANCE)
-cPuellaSAY <- 
-lSponsaSAY <- 
-lQuadrimaculataSAY <- 
-sStriolatumSAY <-
+iElegansSAY <- dplyr::rename(iElegansAY, iElegans_abundance = ABUNDANCE)
+cPuellaSAY <- dplyr::rename(cPuellaAY, cPuella_abundance = ABUNDANCE)
+lSponsaSAY <- dplyr::rename(lSponsaAY, lSponsa_abundance = ABUNDANCE)
+lQuadrimaculataSAY <- dplyr::rename(lQuadrimaculataAY, lQuadrimaculata_abundance = ABUNDANCE)
+sStriolatumSAY <- dplyr::rename(sStriolatumAY, sStriolatum_abundance = ABUNDANCE)
 
@@
 
-dragonflyYears <- c(1959:1988)
+YEAR <- c(1959:1988)
 
@@
 
-bPratenseCompAY <- data.frame(bPratenseAbundance, dragonflyYears)
-pNymphulaCompAY <- data.frame(pNymphulaAbundance, dragonflyYears)
-aCyaneaCompAY <- data.frame(aCyaneaAbundance, dragonflyYears)
-sSanguineumCompAY <- data.frame(sSanguineumAbundance, dragonflyYears)
-aGrandisCompAY <- data.frame(aGrandisAbundance, dragonflyYears)
+bPratenseCompAY <- data.frame(bPratenseAbundance, YEAR)
+pNymphulaCompAY <- data.frame(pNymphulaAbundance, YEAR)
+aCyaneaCompAY <- data.frame(aCyaneaAbundance, YEAR)
+sSanguineumCompAY <- data.frame(sSanguineumAbundance, YEAR)
+aGrandisCompAY <- data.frame(aGrandisAbundance, YEAR)
 
-# Damselfly Larval Data
+# I am going to join these 10 dataframes by the common column year
+# I want year on the far left so I am going to reorder the first dataframe that will determine the order of my join function
 
-larval_data <- read.csv("~/GitHub/rrirfinalproject/larval_data.csv")
+bPratenseCompAY2 <- subset(bPratenseCompAY, select = c(2,1))
+
+join1 <- dplyr::left_join(bPratenseCompAY2, pNymphulaCompAY, by= NULL)
+join2 <- dplyr::left_join(join1, aCyaneaCompAY, by = NULL)
+join3 <- dplyr::left_join(join2, sSanguineumCompAY, by = NULL)
+join4 <- dplyr::left_join(join3, aGrandisCompAY, by = NULL)
+join5 <- dplyr::left_join(join4, cPuellaCompAY, by = NULL)
+join6 <- dplyr::left_join(join5, iElegansCompAY, by = NULL)
+join7 <- dplyr::left_join(join6, lSponsaCompAY, by = NULL)
+join8 <- dplyr::left_join(join7, lQuadrimaculataCompAY, by = NULL)
+fullAbundanceDF <- dplyr::left_join(join8, sStriolatumCompAY, by = NULL)
+
+# I want a total for each year
+
+abundanceSums <- data.frame(rowSums(fullAbundanceDF[,c(2,3,4,5,6,7,8,9,10,11)]))

### 2020-11-25 analysis/analysisSS.Rmd
working on plots

@@
 
-iElegansSAY <- dplyr::rename(iElegansAY, iElegans_abundance = ABUNDANCE)
-cPuellaSAY <- dplyr::rename(cPuellaAY, cPuella_abundance = ABUNDANCE)
-lSponsaSAY <- dplyr::rename(lSponsaAY, lSponsa_abundance = ABUNDANCE)
-lQuadrimaculataSAY <- dplyr::rename(lQuadrimaculataAY, lQuadrimaculata_abundance = ABUNDANCE)
-sStriolatumSAY <- dplyr::rename(sStriolatumAY, sStriolatum_abundance = ABUNDANCE)
+iElegansSAY <- dplyr::rename(iElegansAY, iElegansAbundance = ABUNDANCE)
+cPuellaSAY <- dplyr::rename(cPuellaAY, cPuellaAbundance = ABUNDANCE)
+lSponsaSAY <- dplyr::rename(lSponsaAY, lSponsaAbundance = ABUNDANCE)
+lQuadrimaculataSAY <- dplyr::rename(lQuadrimaculataAY, lQuadrimaculataAbundance = ABUNDANCE)
+sStriolatumSAY <- dplyr::rename(sStriolatumAY, sStriolatumAbundance = ABUNDANCE)
 
@@
 abundanceSums <- data.frame(rowSums(fullAbundanceDF[,c(2,3,4,5,6,7,8,9,10,11)]))
+abundance_year <- cbind(abundanceSums, new_col = YEAR)
+abundanceYearSums <- subset(abundance_year, select = c(2,1))
+
+abunYearSums_1 <- dplyr::rename(abundanceYearSums, year = new_col)
+abunYearsSums <- dplyr::rename(abunYearSums_1, total_abundance = rowSums.fullAbundanceDF...c.2..3..4..5..6..7..8..9..10..11...)
+
+plot(abunYearsSums$year, abunYearsSums$total_abundance)
+
+# I want plots for all 10 species
+
+par(mfrow=c(2,5))
+plot(bPratenseCompAY$YEAR, bPratenseCompAY$bPratenseAbundance, main = "Brachytron pratense", xlab = "Year", ylab = "Abundance")
+plot(pNymphulaCompAY$YEAR, pNymphulaCompAY$pNymphulaAbundance, main = "Pyrrhosa nymphula", xlab = "Year", ylab = "Abundance")
+plot(aCyaneaCompAY$YEAR, aCyaneaCompAY$aCyaneaAbundance, main = "Aeshna cyanea", xlab = "Year", ylab = "Abundance")
+plot(sSanguineumCompAY$YEAR, sSanguineumCompAY$sSanguineumAbundance, main = "Sympetrum sanguineum", xlab = "Year", ylab = "Abundance")
+plot(aGrandisCompAY$YEAR, aGrandisCompAY$aGrandisAbundance, main = "Aeshna grandis", xlab = "Year", ylab = "Abundance")
+plot(cPuellaCompAY$YEAR, cPuellaCompAY$cPuellaAbundance, main = "Coenagrion puella", xlab = "Year", ylab = "Abundance")
+plot(iElegansCompAY$YEAR, iElegansCompAY$iElegansAbundance, main = "Ischnura elegans", xlab = "Year", ylab = "Abundance")
+plot(lSponsaCompAY$YEAR, lSponsaCompAY$lSponsaAbundance, main = "Lestes sponsa", xlab = "Year", ylab = "Abundance")
+plot(lQuadrimaculataCompAY$YEAR, lQuadrimaculataCompAY$lQuadrimaculataAbundance, main = "Libellula quadrimaculata", xlab = "Year", ylab = "Abundance")
+plot(sStriolatumCompAY$YEAR, sStriolatumCompAY$sStriolatumAbundance, main = "Sympetrum striolatum", xlab = "Year", ylab = "Abundance")
+
+# read in Odonate names, this is a .csv in the github repo
+
+odonata_names <- read.csv("~/GitHub/rrirfinalproject/odonata_names.csv")

### 2020-11-26 analysis/analysisSS.Rmd
plotting

@@
 plot(bPratenseCompAY$YEAR, bPratenseCompAY$bPratenseAbundance, main = "Brachytron pratense", xlab = "Year", ylab = "Abundance")
+abline(lm (bPratenseCompAY$bPratenseAbundance ~ bPratenseCompAY$YEAR), col="red")
+
 plot(pNymphulaCompAY$YEAR, pNymphulaCompAY$pNymphulaAbundance, main = "Pyrrhosa nymphula", xlab = "Year", ylab = "Abundance")
+abline(lm (pNymphulaCompAY$pNymphulaAbundance ~ pNymphulaCompAY$YEAR), col="red")
+
 plot(aCyaneaCompAY$YEAR, aCyaneaCompAY$aCyaneaAbundance, main = "Aeshna cyanea", xlab = "Year", ylab = "Abundance")
+abline(lm(aCyaneaCompAY$aCyaneaAbundance ~ aCyaneaCompAY$YEAR), col="red")
+
 plot(sSanguineumCompAY$YEAR, sSanguineumCompAY$sSanguineumAbundance, main = "Sympetrum sanguineum", xlab = "Year", ylab = "Abundance")
+abline(lm(sSanguineumCompAY$sSanguineumAbundance ~ sSanguineumCompAY$YEAR), col="red")
+
 plot(aGrandisCompAY$YEAR, aGrandisCompAY$aGrandisAbundance, main = "Aeshna grandis", xlab = "Year", ylab = "Abundance")
+abline(lm(aGrandisCompAY$aGrandisAbundance ~ aGrandisCompAY$YEAR), col="red")
+
 plot(cPuellaCompAY$YEAR, cPuellaCompAY$cPuellaAbundance, main = "Coenagrion puella", xlab = "Year", ylab = "Abundance")
+abline(lm(cPuellaCompAY$cPuellaAbundance ~ cPuellaCompAY$YEAR), col="red")
+
 plot(iElegansCompAY$YEAR, iElegansCompAY$iElegansAbundance, main = "Ischnura elegans", xlab = "Year", ylab = "Abundance")
+abline(lm(iElegansCompAY$iElegansAbundance ~ iElegansCompAY$YEAR), col="red")
+
 plot(lSponsaCompAY$YEAR, lSponsaCompAY$lSponsaAbundance, main = "Lestes sponsa", xlab = "Year", ylab = "Abundance")
+abline(lm(lSponsaCompAY$lSponsaAbundance ~ lSponsaCompAY$YEAR), col="red")
+
 plot(lQuadrimaculataCompAY$YEAR, lQuadrimaculataCompAY$lQuadrimaculataAbundance, main = "Libellula quadrimaculata", xlab = "Year", ylab = "Abundance")
+abline(lm(lQuadrimaculataCompAY$lQuadrimaculataAbundance ~ lQuadrimaculataCompAY$YEAR), col="red")
+
 plot(sStriolatumCompAY$YEAR, sStriolatumCompAY$sStriolatumAbundance, main = "Sympetrum striolatum", xlab = "Year", ylab = "Abundance")
+abline(lm(sStriolatumCompAY$sStriolatumAbundance ~ sStriolatumCompAY$YEAR), col="red")
+
+fitBPratense <- lm(bPratenseCompAY$bPratenseAbundance ~ bPratenseCompAY$YEAR)
 

### 2020-11-29 analysis/analysisSS.Rmd
working on plots, had to rename data file

@@
 
+DataSingleStudy <- read.csv("~/GitHub/rrirfinalproject/DataSingleStudy")
 rawDragonfly <- `DataSingleStudy`
@@
 
-plot(abunYearsSums$year, abunYearsSums$total_abundance)
+par(mfrow=c(1,1))
+plot(abunYearsSums$year, abunYearsSums$total_abundance, xlab = "Year", ylab = "Abundance")
+abline(lm(abunYearsSums$total_abundance ~ abunYearsSums$year), col="red")
 
-# I want plots for all 10 species
+fitTotal <- lm(abunYearsSums$total_abundance ~ abunYearsSums$year)
+
+summary(fitTotal)
+
+pValueTotal <- 0.000162
+
+# I want plots for all 10 species with a trend line
 
@@
 
+# I want to see if these are significant, so I am going to save the lm and then I will do a summary of each one
+
 fitBPratense <- lm(bPratenseCompAY$bPratenseAbundance ~ bPratenseCompAY$YEAR)
+summary(fitBPratense)
+pValueBPratense <- 0.5911
+
+fitPNymphula <- lm(pNymphulaCompAY$pNymphulaAbundance ~ pNymphulaCompAY$YEAR)
+summary(fitPNymphula)
+pValuePNymphula <- 0.02333
+
+fitACyanea <- lm(aCyaneaCompAY$aCyaneaAbundance ~ aCyaneaCompAY$YEAR)
+summary(fitACyanea)
+pValueACyanea <- 0.3033
+
+fitSSanguineum <- lm(sSanguineumCompAY$sSanguineumAbundance ~ sSanguineumCompAY$YEAR)
+summary(fitSSanguineum)
+pValueSSanguineum <- 0.00009631
+
+fitAGrandis <- lm(aGrandisCompAY$aGrandisAbundance ~ aGrandisCompAY$YEAR)
+summary(fitAGrandis)
+pValueAGrandis <- 0.1363
+
+fitCPuella <- lm(cPuellaCompAY$cPuellaAbundance ~ cPuellaCompAY$YEAR)
+summary(fitCPuella)
+pValueCPuella <- 8.362e-07
+
+fitIElegans <- lm(iElegansCompAY$iElegansAbundance ~ iElegansCompAY$YEAR)
+summary(fitIElegans)
+pValueIElegans <- 0.103
+
+fitLSponsa <- lm(lSponsaCompAY$lSponsaAbundance ~ lSponsaCompAY$YEAR)
+summary(fitLSponsa)
+pValueLSponsa <- 0.0001868
+
+fitLQuadrimaculata <- lm(lQuadrimaculataCompAY$lQuadrimaculataAbundance ~ lQuadrimaculataCompAY$YEAR)
+summary(fitLQuadrimaculata)
+pValueLQuadrimaculata <- 0.0359
+
+fitSStriolatum <- lm(sStriolatumCompAY$sStriolatumAbundance ~ sStriolatumCompAY$YEAR)
+summary(fitSStriolatum)
+pValueSStriolatum <- 0.9176
 

### 2020-12-01 analysis/analysisSS.Rmd
tweaking species numbers

@@
 
+# I want to know how many species there are in a given year, using the nrow function from base
+# I will save each as with a corresponding year identifier because I have not yet decided what format I want them all in (dataframe, table, etc.)
+
+ASpecies <- length(which(fullAbundanceDF[1,2:11] != 0))
+BSpecies <- length(which(fullAbundanceDF[2,2:11] != 0))
+CSpecies <- length(which(fullAbundanceDF[3,2:11] != 0))
+DSpecies <- length(which(fullAbundanceDF[4,2:11] != 0))
+ESpecies <- length(which(fullAbundanceDF[5,2:11] != 0))
+FSpecies <- length(which(fullAbundanceDF[6,2:11] != 0))
+GSpecies <- length(which(fullAbundanceDF[7,2:11] != 0))
+HSpecies <- length(which(fullAbundanceDF[8,2:11] != 0))
+ISpecies <- length(which(fullAbundanceDF[9,2:11] != 0))
+JSpecies <- length(which(fullAbundanceDF[10,2:11] != 0))
+KSpecies <- length(which(fullAbundanceDF[11,2:11] != 0))
+LSpecies <- length(which(fullAbundanceDF[12,2:11] != 0))
+MSpecies <- length(which(fullAbundanceDF[13,2:11] != 0))
+NSpecies <- length(which(fullAbundanceDF[14,2:11] != 0))
+OSpecies <- length(which(fullAbundanceDF[15,2:11] != 0))
+PSpecies <- length(which(fullAbundanceDF[16,2:11] != 0))
+QSpecies <- length(which(fullAbundanceDF[17,2:11] != 0))
+RSpecies <- length(which(fullAbundanceDF[18,2:11] != 0))
+SSpecies <- length(which(fullAbundanceDF[19,2:11] != 0))
+TSpecies <- length(which(fullAbundanceDF[20,2:11] != 0))
+USpecies <- length(which(fullAbundanceDF[21,2:11] != 0))
+VSpecies <- length(which(fullAbundanceDF[22,2:11] != 0))
+XSpecies <- length(which(fullAbundanceDF[23,2:11] != 0))
+YSpecies <- length(which(fullAbundanceDF[24,2:11] != 0))
+ZSpecies <- length(which(fullAbundanceDF[25,2:11] != 0))
+AASpecies <- length(which(fullAbundanceDF[26,2:11] != 0))
+ABSpecies <- length(which(fullAbundanceDF[27,2:11] != 0))
+ACSpecies <- length(which(fullAbundanceDF[28,2:11] != 0))
+ADSpecies <- length(which(fullAbundanceDF[29,2:11] != 0))
+AESpecies <- length(which(fullAbundanceDF[30,2:11] != 0))
+
 # read in Odonate names, this is a .csv in the github repo
 
-odonata_names <- read.csv("~/GitHub/rrirfinalproject/odonata_names.csv")
+odonata_names <- read.table("~/GitHub/rrirfinalproject/odonata_names.csv")

### 2020-12-01 manuscript/manuscript.Rmd
manuscript rmd added, notes previously taken compiled

---
title: "Odonate abundance changes in response to the creation of a flood reservoir in a former fen"
author: Maggie Vincent
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The invertebrate order Odonata is comprised of at least 5680 freshwater insects and are commonly referred to as damselflies and dragonflies (Kalkman et al. 2008).

The order is of particular interest for their reproductive behaviors, with males of the species utilizing the aedeagus not only for sperm transfer, but also for the removal of sperm already in a female’s storage organs (Waage, 1979).

Destruction of habitation and changes in water quality have led to decreases in species distributions, although extinction is not an immediate worry for many (Kalkman et al. 2008).

Although the highest biodiversity of Odonates resides in the Neotropics and where? (Citation), Odonates extend into temperate zones, albeit with less diversification. This project is concerned with East Anglia, a region in eastern England. This area was formerly swampy, or fenland, until it was drained for agricultural use over the last few centuries (DeFries et al. 2004). Habitat changes such as fragmentation and/or loss often lead to decreases in biodiversity (DeFries et al. 2004), as stated above for Odonates. Conservation and preservation attempts over the last century have moved to protect remaining areas and return others back to their original (or something similar). There has been work since the late 1950s in East Anglia to preserve ancient fens with the possibility of reconnecting the remaining fragments into a “Great Fen” (Citation).

# Objectives

The goals of this study can be separated into two main purposes: description of changing (or not) species diversity over more than two decades and a focus on transparency and reproducibility in the workflow.

## Data Source

The data for this descriptive study is available through BioTIME (biotime.st-andrews.ac.uk/downloadArea.php), a database devoted to biodiversity. The database is dynamic and does not have an API, making reproducibility via the database a step more difficult. The dataset itself is constant, but its online source changing. The dataset was not originally representative of the first study’s findings, and required significant manipulation. The methods and resulting implications for reproducibility will be discussed in more detail in later sections.

[Head of original data set?] – head()

## Study Site

The work was carried out at Woodwalton Fen (Long, Lat) in Huntingdon, England, from 1959 to 1988. The area was primarily an ancient fen, or a wetland, until most of the area was drained in the 1800s for agricultural use. In the mid-1900s the study area was restructured as a flood reservoir, recreating the wetland aspect. The original study followed Odonata species to characterize territorial changes and patterns throughout the project (Citation).

[Images of Woodwalton Fen]

## Approach

The original study characterized the return and territoriality of dragonflies and damselflies at the site. In this study, I seek to better characterize the significance of species-specific changes over the study period in order to tease apart any possible changes in overall abundance. The main questions can be summarized as follows:

# Methods

The data used is publicly available for download and use from BioTIME (citation) or from the Github repository hosting the analytical pipeline for this project. The data was manipulated (with an original retained) in R Version 4.0.2 with the additional packages ‘plyr’ and ‘dplyr’. The manipulated dataset is also available on the aforementioned Github repository, along with the code used in the analysis folder. As the dataset was previously compiled and published, the methods for this project will concern the approach taken and code utilized.

Base, plyr and dplyr were utilized in an attempt to keep the number of packages small and the threshold for reproducibility lower. The latter two aided in the manipulation of the dataset.

# Results

# Discussion

# Reproducibility

The unmanipulated dataset and analysis are hosted on a Github repository (link), along with the manipulated dataset and corresponding code, figures, license information and a Docker file. The dataset was not called for using an API, making it necessary to download externally and import into R. The original directions are documented in the analysis (on Github) and the unmanipulated dataset can also be downloaded from the Github repository. As mentioned earlier, the original dataset had some issues that had to be resolved prior to analysis. Abundances of “zero” were not entered into the dataset, leaving that year out of the dataset entirely. Once these were returned the dataset, further manipulation and analysis was possible. The analysis, an R markdown file, details how to download the dataset and shows the code for manipulation in order to minimize any reproducibility issues and show transparency.

# Conclusion

# References

Kalkman, V. J., Clausnitzer V., Dijkstra K. B., Orr A. G., Paulson D. R., van Tol J. (2008) Global diversity of dragonflies (Odonata) in freshwater. Hydrobiologia, 595, 351-363.

Waage, J. K. (1979) Dual function of the damselfly penis: Sperm removal and transfer. Science, 203:4383, 916-918.

### 2020-12-01 manuscript/manuscript.Rmd
added picture; worked on manuscript, spec. the intro, obj., methods, results and concl.

@@
 author: Maggie Vincent
-output: pdf_document
+output:
+  pdf_document:
+    fig_caption: yes
+
 ---
@@
 
+```{r, echo=FALSE}
+install.packages('plyr')
+install.packages('dplyr')
+```
+
 # Introduction
 
-The invertebrate order Odonata is comprised of at least 5680 freshwater insects and are commonly referred to as damselflies and dragonflies (Kalkman et al. 2008).
+The invertebrate order Odonata is comprised of at least 5680 freshwater insects and are commonly referred to as damselflies and dragonflies.
+Due to the larvae’s need for freshwater, they can be used as gauges for local environmental quality and health (Kalkman *et al.* 2008).
+Destruction of habitation and changes in water quality have led to decreases in species distributions, although extinction is not an immediate worry for most, as they often have broad distributions (Kalkman *et al.* 2008).
 
-The order is of particular interest for their reproductive behaviors, with males of the species utilizing the aedeagus not only for sperm transfer, but also for the removal of sperm already in a female’s storage organs (Waage, 1979).
+The order is of particular interest for their reproductive behaviors, with males of the species utilizing the aedeagus not only for sperm transfer from a secondary sexual organ, but also for the removal of sperm already in a female’s storage organs (Waage, 1979).
+Additionally, because many species are quite conspicuous in their coloration, they are possible for identification by non-experts.
+This makes them an interesting and entertaining aspect of outdoor recreation and wildlife hobbyists.
 
-Destruction of habitation and changes in water quality have led to decreases in species distributions, although extinction is not an immediate worry for many (Kalkman et al. 2008).
+Although the highest biodiversity of Odonates resides in the Neotropical and Oriental regions (Kalkman *et al.* 2008), Odonates extend into temperate zones, albeit with less diversification.
+This project is concerned with East Anglia, a region in eastern England.
+This area was formerly swampy, or fenland, until it was drained for agricultural use over the last few centuries (DeFries *et al.* 2004).
+Habitat changes such as fragmentation and/or loss often lead to decreases in biodiversity (DeFries *et al.* 2004), as stated above for Odonates.
+Conservation and preservation attempts over the last century have moved to protect remaining areas and return others back to their original (or similar) functions.
+There has been work since the late 1950s in East Anglia to preserve ancient fens with the possibility of reconnecting the remaining fragments into a “Great Fen” (“Great”, n.d.) in the relevant area.
 
-Although the highest biodiversity of Odonates resides in the Neotropics and where?
-(Citation), Odonates extend into temperate zones, albeit with less diversification.
-This project is concerned with East Anglia, a region in eastern England.
-This area was formerly swampy, or fenland, until it was drained for agricultural use over the last few centuries (DeFries et al.
-2004).
-Habitat changes such as fragmentation and/or loss often lead to decreases in biodiversity (DeFries et al.
-2004), as stated above for Odonates.
-Conservation and preservation attempts over the last century have moved to protect remaining areas and return others back to their original (or something similar).
-There has been work since the late 1950s in East Anglia to preserve ancient fens with the possibility of reconnecting the remaining fragments into a “Great Fen” (Citation).
+The recent movement to assess the environmental-agricultural trade-offs of reclaimed lands in England (and elsewhere) requires a careful balance between the environmental benefits of conservation and the financial difficulties for farmers and locals (Morris et al.
+2000).
+There is the possibility of determining areas of lower agricultural production that can be returned to their former state without detrimental effects to local famers (Morris *et al.* 2000; “Great”, n.d.).
 
@@
 
-The goals of this study can be separated into two main purposes: description of changing (or not) species diversity over more than two decades and a focus on transparency and reproducibility in the workflow.
+The goals of this study can be separated into two main purposes: description of changing (or not) species abundance over more than two decades and a focus on transparency and reproducibility in the workflow.
+
+* Does *Odonata* abundance significantly change over the course of the study?
+* Do species appear to follow different patterns of abundance changes?
+* Do certain species seem to suffer upon the introduction of others?
+  + How does this match up with current citizen science?
+
+The first three main objectives will be answered from the dataset from the original study (Morris, 1991).
+The final sub-bullet is not meant to be answered with true rigor, but as an interesting way to include citizen science.
+It is not expected to be comprehensive or a full picture of current Odonate diversity at the study site.
 
@@
 
-[Head of original data set?] – head()
+```{r, echo=FALSE, fig.cap = 'Sample of the unmanipulated data set.'}
+DataSingleStudy <- read.csv("~/GitHub/rrirfinalproject/DataSingleStudy")
+head(DataSingleStudy)
+```
 
@@
 
-[Images of Woodwalton Fen]
+![Map of eastern England with Woodwalton Fen marked.](Woodwalton_Fen_England.jpg)
 
@@
 In this study, I seek to better characterize the significance of species-specific changes over the study period in order to tease apart any possible changes in overall abundance.
-The main questions can be summarized as follows:
+The main questions were outlined earlier in the objectives.
 
@@
 
-The data used is publicly available for download and use from BioTIME (citation) or from the Github repository hosting the analytical pipeline for this project.
+The data used is publicly available for download and use from BioTIME (Dornela *et al.* 2018) or from the Github repository (https://github.com/mvcarneal/rrirfinalproject) hosting the analytical pipeline for this project.
 The data was manipulated (with an original retained) in R Version 4.0.2 with the additional packages ‘plyr’ and ‘dplyr’.
@@
 
-Base, plyr and dplyr were utilized in an attempt to keep the number of packages small and the threshold for reproducibility lower. The latter two aided in the manipulation of the dataset.
+Base, plyr and dplyr were utilized in an attempt to keep the number of packages small and the threshold for reproducibility lower, and Docker used to create a container for reproducibility.
+The latter two aided in the manipulation of the dataset.
 
 # Results
+
+```{r, echo=FALSE}
+library('plyr')
+library('dplyr')
+
+DataSingleStudy <- read.csv("~/GitHub/rrirfinalproject/DataSingleStudy")
+rawDragonfly <- `DataSingleStudy`
+
+cPuella <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Coenagrion puella')
+iElegans <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Ischnura elegans')
+lSponsa <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Lestes sponsa')
+lQuadrimaculata <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Libellula quadrimaculata')
+sStriolatum <- dplyr::filter(rawDragonfly, GENUS_SPECIES == 'Sympetrum striolatum')
+
+getAbundanceYear <- function(dat, x = 'ABUNDANCE', y = 'YEAR'){
+                    tmp <- subset(dat, select = c(x, y))
+                    return(tmp)
+}
+
+iElegansAY <- getAbundanceYear(iElegans)
+cPuellaAY <- getAbundanceYear(cPuella)
+lSponsaAY <- getAbundanceYear(lSponsa)
+lQuadrimaculataAY <- getAbundanceYear(lQuadrimaculata)
+sStriolatumAY <- getAbundanceYear(sStriolatum)
+
+iElegansSAY <- dplyr::rename(iElegansAY, iElegansAbundance = ABUNDANCE)
+cPuellaSAY <- dplyr::rename(cPuellaAY, cPuellaAbundance = ABUNDANCE)
+lSponsaSAY <- dplyr::rename(lSponsaAY, lSponsaAbundance = ABUNDANCE)
+lQuadrimaculataSAY <- dplyr::rename(lQuadrimaculataAY, lQuadrimaculataAbundance = ABUNDANCE)
+sStriolatumSAY <- dplyr::rename(sStriolatumAY, sStriolatumAbundance = ABUNDANCE)
+
+row1959 <- c(0, 1959)
+row1960 <- c(0, 1960)
+row1961 <- c(0, 1961)
+row1962 <- c(0, 1962)
+row1963 <- c(0, 1963)
+
+row1972 <- c(0, 1972)
+row1974 <- c(0, 1974)
+
+cPuellaCompAY <- rbind(row1959, row1960, row1961, row1962, cPuellaSAY)
+iElegansCompAY <- rbind(row1959, row1960, row1961, iElegansSAY)
+lSponsaCompAY <- rbind(row1959, row1960, row1961, row1962, row1963, lSponsaSAY)
+# lQuadrimaculata will be dealt with differently because it is missing rows within the dataframe
+sStriolatumCompAY <- rbind(row1959, row1960, row1961, sStriolatumSAY)
+
+prelibellulaCompAY <- rbind(row1960, row1972, row1974, lQuadrimaculataSAY)
+# this will add the rows I want, but now in chronological order
+lQuadrimaculataCompAY <- dplyr::arrange(prelibellulaCompAY, YEAR)
+
+# there are species missing from this dataset, they need to be added separately, I will source (and cite) the abundance data
+# also need a vector corresponding to the years
+
+YEAR <- c(1959:1988)
+
+bPratenseAbundance <- c(0,0,0,0,0,0,0,0,0,1,0,0,2,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
+pNymphulaAbundance <- c(0,0,3,20,18,5,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
+aCyaneaAbundance <- c(0,0,0,0,1,2,3,1,1,2,3,1,5,1,2,0,0,0,1,1,1,0,2,1,1,3,1,2,2,2)
+sSanguineumAbundance <- c(0,0,0,0,0,1,0,0,1,0,2,0,1,0,0,1,0,1,4, 11, 3, 0, 1, 9, 6, 10, 4, 9, 3, 4)
+aGrandisAbundance <- c(0,0,0,0,0,0,0,4,0,1,1,0,2,0,2,0,0,0,1,1,0,1,2,1,0,1,1,4,0,1)
+
+bPratenseCompAY <- data.frame(bPratenseAbundance, YEAR)
+pNymphulaCompAY <- data.frame(pNymphulaAbundance, YEAR)
+aCyaneaCompAY <- data.frame(aCyaneaAbundance, YEAR)
+sSanguineumCompAY <- data.frame(sSanguineumAbundance, YEAR)
+aGrandisCompAY <- data.frame(aGrandisAbundance, YEAR)
+
+# I am going to join these 10 dataframes by the common column year
+# I want year on the far left so I am going to reorder the first dataframe that will determine the order of my join function
+
+bPratenseCompAY2 <- subset(bPratenseCompAY, select = c(2,1))
+
+join1 <- dplyr::left_join(bPratenseCompAY2, pNymphulaCompAY, by= NULL)
+join2 <- dplyr::left_join(join1, aCyaneaCompAY, by = NULL)
+join3 <- dplyr::left_join(join2, sSanguineumCompAY, by = NULL)
+join4 <- dplyr::left_join(join3, aGrandisCompAY, by = NULL)
+join5 <- dplyr::left_join(join4, cPuellaCompAY, by = NULL)
+join6 <- dplyr::left_join(join5, iElegansCompAY, by = NULL)
+join7 <- dplyr::left_join(join6, lSponsaCompAY, by = NULL)
+join8 <- dplyr::left_join(join7, lQuadrimaculataCompAY, by = NULL)
+fullAbundanceDF <- dplyr::left_join(join8, sStriolatumCompAY, by = NULL)
+
+# I want a total for each year
+
+abundanceSums <- data.frame(rowSums(fullAbundanceDF[,c(2,3,4,5,6,7,8,9,10,11)]))
+abundance_year <- cbind(abundanceSums, new_col = YEAR)
+abundanceYearSums <- subset(abundance_year, select = c(2,1))
+
+abunYearSums_1 <- dplyr::rename(abundanceYearSums, year = new_col)
+abunYearsSums <- dplyr::rename(abunYearSums_1, total_abundance = rowSums.fullAbundanceDF...c.2..3..4..5..6..7..8..9..10..11...)
+
+par(mfrow=c(1,1))
+plot(abunYearsSums$year, abunYearsSums$total_abundance, xlab = "Year", ylab = "Abundance")
+abline(lm(abunYearsSums$total_abundance ~ abunYearsSums$year), col="red")
+
+fitTotal <- lm(abunYearsSums$total_abundance ~ abunYearsSums$year)
+
+summary(fitTotal)
+
+pValueTotal <- 0.000162
+
+```
 
@@
 
-The unmanipulated dataset and analysis are hosted on a Github repository (link), along with the manipulated dataset and corresponding code, figures, license information and a Docker file.
+The unmanipulated dataset and analysis are hosted on a Github repository (linked in the Methods section), along with the manipulated dataset and corresponding code, figures, license information and a Docker file.
 The dataset was not called for using an API, making it necessary to download externally and import into R.
@@
 Abundances of “zero” were not entered into the dataset, leaving that year out of the dataset entirely.
+Several species were also left out, but using the orignal study were returned to the data set.
+An image from the original study is in the analysis folder on the Github repository for easy reference (Morris, 1991).
 Once these were returned the dataset, further manipulation and analysis was possible.
@@
 
-# Conclusion
+# Conclusions
+
+This project was inherently limited by the structure and lack of completeness of the data set obtained.
+Perhaps the most important conclusion of this project is the forethought required in how data is collected and databased for future use.
+Morris, the original researcher, collected this information in a time when these considerations did not need to be taken, as open source data sets were certainly not on the average researcher's mind.
+While this study is limited in its analytical ability, the author does not intend to masquerade the results as such.
+Much of the ecological information needed in order to determine the most important factors in *Odonata* diversity (vegetation, pond area, scrub management) is not available.
+Therefore, the findings of the original study cannot be expanded upon in an ecological sense.
+It also showcases the difficulty in finding representative and complete data sets, with the admission that this data set might not have been the best choice, at least not on its own.
 
@@
 
-Kalkman, V. J., Clausnitzer V., Dijkstra K. B., Orr A. G., Paulson D. R., van Tol J. (2008) Global diversity of dragonflies (Odonata) in freshwater. Hydrobiologia, 595, 351-363.
+Dornela M, Antao LH, Moyes F, Bates A. E., *et al.* BioTIME: A database of biodiversity time series for the Anthropocene. *Global Ecol Biogeogr.* 2018; 27:760-786. https://doi.org/10.1111/geb.12729
 
-Waage, J. K. (1979) Dual function of the damselfly penis: Sperm removal and transfer. Science, 203:4383, 916-918.
+Kalkman, V. J., Clausnitzer V., Dijkstra K. B., Orr A. G., Paulson D. R., van Tol J. (2008) Global diversity of dragonflies (Odonata) in freshwater. *Hydrobiologia*, **595**, 351-363.
+
+Moore, N.
+W.
+(1991) The development of dragonfly communities and the consequences of territorial behavior: a 27 year study on small ponds at Woodwalton Fen, Cambridgeshire, United Kingdom.
+*Odonatologica*, **20**(2), 203-231.
+
+Waage, J. K. (1979) Dual function of the damselfly penis: Sperm removal and transfer. *Science*, **203**:4383, 916-918.
+
+The Wildlife Trusts (n.d.) About the Great Fen. https://www.greatfen.org.uk/about-great-fen

### 2020-12-02 analysis/analysisSS.Rmd
figures exported from analysis, figures file added for clarity

@@
 
+# number of species in chronological order
+# I'll reuse the years vector I created earlier to make a data.frame
+
+speciesChrono <- c(ASpecies, BSpecies, CSpecies, DSpecies, ESpecies, FSpecies, GSpecies, HSpecies, ISpecies, JSpecies, KSpecies, LSpecies, MSpecies, NSpecies, OSpecies, PSpecies, QSpecies, RSpecies, SSpecies, TSpecies, USpecies, VSpecies, XSpecies, YSpecies, ZSpecies, AASpecies, ABSpecies, ACSpecies, ADSpecies, AESpecies)
+
+speciesByYear <- data.frame(YEAR, speciesChrono)
+
+# look at it graphically
+
+par(mfrow=c(1,1))
+plot(speciesByYear$YEAR, speciesByYear$speciesChrono, xlab = "Year", ylab = "Species Number")
+
 # read in Odonate names, this is a .csv in the github repo

### 2020-12-02 analysis/analysisSS.Rmd
working on results and figures in manuscript

@@
 
-# I want to know how many species there are in a given year, using the nrow function from base
+# I want to know how many species there are in a given year, subsetting using the which function and finding the length of that subset
 # I will save each as with a corresponding year identifier because I have not yet decided what format I want them all in (dataframe, table, etc.)
@@
 odonata_names <- read.table("~/GitHub/rrirfinalproject/odonata_names.csv")
+
+# weird character in genus column name, need to rename it, dplyr function
+
+odonata_Woodwalton <- dplyr::rename(odonata_names, GENUS = ï..GENUS)

### 2020-12-02 manuscript/manuscript.Rmd
working on results and figures in manuscript

@@
   pdf_document:
-    fig_caption: yes
+    fig_caption: true
 
@@
 
-![Map of eastern England with Woodwalton Fen marked.](Woodwalton_Fen_England.jpg)
+![**Figure 1.** Map of eastern England with Woodwalton Fen marked.](Woodwalton_Fen_England.jpg)
 
@@
 abunYearsSums <- dplyr::rename(abunYearSums_1, total_abundance = rowSums.fullAbundanceDF...c.2..3..4..5..6..7..8..9..10..11...)
+```
 
+The damselflies and dragonflies in this study are listed in Table 1 for reference.
+The figure below (Fig.
+2), is a linear regression model of the maximum abundance of males in a single day from 1959  to 1988.
+The abundance corresponds to the sum of the ten species in Table 1 and the p-value for the model was 0.000162.
+
+```{r, echo=FALSE,}
+
+# weird character in genus column name, need to rename it, dplyr function
+
+odonata_Woodwalton <- dplyr::rename(odonata_names, GENUS = ï..GENUS)
+
+knitr::kable(odonata_Woodwalton, caption = 'Table 1. Species and common names of ten Odonates found at Woodwalton Fen between 1959 and 1988.')
+
+```
+
+```{r, echo=FALSE, fig.align=center, fig.cap='**Figure 2** Linear model of abundance of all ten species from 1959 to 1988. p-value = 0.000162 }
 par(mfrow=c(1,1))
@@
 abline(lm(abunYearsSums$total_abundance ~ abunYearsSums$year), col="red")
+```
 
+```{r, echo=TRUE}
 fitTotal <- lm(abunYearsSums$total_abundance ~ abunYearsSums$year)
@@
 pValueTotal <- 0.000162
+
+```
+
+```{r, echo=FALSE, fig.cap=paste("Figure 3. Abundances broken down by species with line of best fit.") }
+par(mfrow=c(2,5))
+plot(bPratenseCompAY$YEAR, bPratenseCompAY$bPratenseAbundance, main = "Brachytron pratense", xlab = "Year", ylab = "Abundance")
+abline(lm (bPratenseCompAY$bPratenseAbundance ~ bPratenseCompAY$YEAR), col="red")
+
+plot(pNymphulaCompAY$YEAR, pNymphulaCompAY$pNymphulaAbundance, main = "Pyrrhosa nymphula", xlab = "Year", ylab = "Abundance")
+abline(lm (pNymphulaCompAY$pNymphulaAbundance ~ pNymphulaCompAY$YEAR), col="red")
+
+plot(aCyaneaCompAY$YEAR, aCyaneaCompAY$aCyaneaAbundance, main = "Aeshna cyanea", xlab = "Year", ylab = "Abundance")
+abline(lm(aCyaneaCompAY$aCyaneaAbundance ~ aCyaneaCompAY$YEAR), col="red")
+
+plot(sSanguineumCompAY$YEAR, sSanguineumCompAY$sSanguineumAbundance, main = "Sympetrum sanguineum", xlab = "Year", ylab = "Abundance")
+abline(lm(sSanguineumCompAY$sSanguineumAbundance ~ sSanguineumCompAY$YEAR), col="red")
+
+plot(aGrandisCompAY$YEAR, aGrandisCompAY$aGrandisAbundance, main = "Aeshna grandis", xlab = "Year", ylab = "Abundance")
+abline(lm(aGrandisCompAY$aGrandisAbundance ~ aGrandisCompAY$YEAR), col="red")
+
+plot(cPuellaCompAY$YEAR, cPuellaCompAY$cPuellaAbundance, main = "Coenagrion puella", xlab = "Year", ylab = "Abundance")
+abline(lm(cPuellaCompAY$cPuellaAbundance ~ cPuellaCompAY$YEAR), col="red")
+
+plot(iElegansCompAY$YEAR, iElegansCompAY$iElegansAbundance, main = "Ischnura elegans", xlab = "Year", ylab = "Abundance")
+abline(lm(iElegansCompAY$iElegansAbundance ~ iElegansCompAY$YEAR), col="red")
+
+plot(lSponsaCompAY$YEAR, lSponsaCompAY$lSponsaAbundance, main = "Lestes sponsa", xlab = "Year", ylab = "Abundance")
+abline(lm(lSponsaCompAY$lSponsaAbundance ~ lSponsaCompAY$YEAR), col="red")
+
+plot(lQuadrimaculataCompAY$YEAR, lQuadrimaculataCompAY$lQuadrimaculataAbundance, main = "Libellula quadrimaculata", xlab = "Year", ylab = "Abundance")
+abline(lm(lQuadrimaculataCompAY$lQuadrimaculataAbundance ~ lQuadrimaculataCompAY$YEAR), col="red")
+
+plot(sStriolatumCompAY$YEAR, sStriolatumCompAY$sStriolatumAbundance, main = "Sympetrum striolatum", xlab = "Year", ylab = "Abundance")
+abline(lm(sStriolatumCompAY$sStriolatumAbundance ~ sStriolatumCompAY$YEAR), col="red")
+
+```
+
+```{r, echo=FALSE}
+library('plyr')
+library('dplyr')
+
+names_p_values <- read.csv("~/GitHub/rrirfinalproject/analysis/names_p_values.csv")
+
+# strange character in column name for species
+
+corrected_names_p_values <- dplyr::rename(names_p_values, Species = ï..SPECIES)
+
+knitr::kable(corrected_names_p_values, caption = 'Table 2. p-values from the LM for each species and all species combined shown in Figure 2. Significance shown with asterisks.')
+
+```
+
+```{r, echo=FALSE, fig.cap="Figure 4.
+Number of species in each year of the study period.
+Note that this is not a comprehensive list of all Odonates at the site, this will be addressed in the discussion."}
+# I want to know how many species there are in a given year, subsetting using the which function and finding the length of that subset
+# I will save each as with a corresponding year identifier because I have not yet decided what format I want them all in (dataframe, table, etc.)
+
+ASpecies <- length(which(fullAbundanceDF[1,2:11] != 0))
+BSpecies <- length(which(fullAbundanceDF[2,2:11] != 0))
+CSpecies <- length(which(fullAbundanceDF[3,2:11] != 0))
+DSpecies <- length(which(fullAbundanceDF[4,2:11] != 0))
+ESpecies <- length(which(fullAbundanceDF[5,2:11] != 0))
+FSpecies <- length(which(fullAbundanceDF[6,2:11] != 0))
+GSpecies <- length(which(fullAbundanceDF[7,2:11] != 0))
+HSpecies <- length(which(fullAbundanceDF[8,2:11] != 0))
+ISpecies <- length(which(fullAbundanceDF[9,2:11] != 0))
+JSpecies <- length(which(fullAbundanceDF[10,2:11] != 0))
+KSpecies <- length(which(fullAbundanceDF[11,2:11] != 0))
+LSpecies <- length(which(fullAbundanceDF[12,2:11] != 0))
+MSpecies <- length(which(fullAbundanceDF[13,2:11] != 0))
+NSpecies <- length(which(fullAbundanceDF[14,2:11] != 0))
+OSpecies <- length(which(fullAbundanceDF[15,2:11] != 0))
+PSpecies <- length(which(fullAbundanceDF[16,2:11] != 0))
+QSpecies <- length(which(fullAbundanceDF[17,2:11] != 0))
+RSpecies <- length(which(fullAbundanceDF[18,2:11] != 0))
+SSpecies <- length(which(fullAbundanceDF[19,2:11] != 0))
+TSpecies <- length(which(fullAbundanceDF[20,2:11] != 0))
+USpecies <- length(which(fullAbundanceDF[21,2:11] != 0))
+VSpecies <- length(which(fullAbundanceDF[22,2:11] != 0))
+XSpecies <- length(which(fullAbundanceDF[23,2:11] != 0))
+YSpecies <- length(which(fullAbundanceDF[24,2:11] != 0))
+ZSpecies <- length(which(fullAbundanceDF[25,2:11] != 0))
+AASpecies <- length(which(fullAbundanceDF[26,2:11] != 0))
+ABSpecies <- length(which(fullAbundanceDF[27,2:11] != 0))
+ACSpecies <- length(which(fullAbundanceDF[28,2:11] != 0))
+ADSpecies <- length(which(fullAbundanceDF[29,2:11] != 0))
+AESpecies <- length(which(fullAbundanceDF[30,2:11] != 0))
+
+# number of species in chronological order
+# I'll reuse the years vector I created earlier to make a data.frame
+
+speciesChrono <- c(ASpecies, BSpecies, CSpecies, DSpecies, ESpecies, FSpecies, GSpecies, HSpecies, ISpecies, JSpecies, KSpecies, LSpecies, MSpecies, NSpecies, OSpecies, PSpecies, QSpecies, RSpecies, SSpecies, TSpecies, USpecies, VSpecies, XSpecies, YSpecies, ZSpecies, AASpecies, ABSpecies, ACSpecies, ADSpecies, AESpecies)
+
+speciesByYear <- data.frame(YEAR, speciesChrono)
+
+# look at it graphically
+
+par(mfrow=c(1,1))
+plot(speciesByYear$YEAR, speciesByYear$speciesChrono, xlab = "Year", ylab = "Species Number")
 

### 2020-12-03 manuscript/manuscript.Rmd
first work on discussion
work on discussion, concl., references

@@
 
+Taken together, the abundance of all ten species had a clear upward trend in abundance from 1959 to 1988 (Fig.
+2).
+When modeled, the change in abundance was significant (Table 2), indicating the pond construction at Woodwalton Fen increased Odonate abundance.
+In order to see if certain species were driving these abundance changes, each species was analyzed with a simple linear regression model (Fig.
+3) and fitted with a line of best fit.
+Of the ten species, five had significant changes in abundance over the study period (Table 2), indicating not all species were responsible for the overall change in abundance.
+Three species had significant increases in abundance, *C.
+puella*, *S.
+Sanguineum*, and *L.
+sponsa*.
+Interestingly, two species had significant downtrends, *L.
+quadrimaculata* started at over 150 individuals and dropped to less than a third of that for the last decade of the study period.
+*P.
+nymphula* was also indicated as a significant abundance decrease in the model, but with a smaller peak abundance (~20), the results should be taken with a grain of salt.
+Of the ten species, six of them had peak abundances of less than 25, meaning errors in field observations could easily sway the findings.
+This should be considered in the species with a significant decrease (*P.
+nymphula*) and possibly *B.
+pratense*.
+The peak abundance for this species was two, with the species not reported after the mid-seventies.
+While not significant for the abundance patterns objectives, it will be interesting when compared later against current citizen science.
+
+Figure 4 shows the number of species per year, corresponding to the final objective concerning species diversity and richness.
+The model shows a period of early, rapid colonization and then a maintenance of eight of the ten species.
+However, upon consultation with the original study (Moore, 1991), it became apparent that the ten species were not the only ones seen at Woodwalton Fen over the study period.
+As such, Figure 4 cannot be considered a comprehensive description of Odonate species diversity at the site.
+Further analysis considering evenness and richness were halted so as to not mischaracterize the site.
+The original study, however, did mention a similar pattern of rapid colonization early on, potentially holding with this study’s findings of a subset of the species.
+
+Due to the time in between the original study and this one (~30 years), current citizen science from iNaturalist (https://www.inaturalist.org/, via GBIf, https://www.gbif.org/) was incorporated merely out of curiosity of the researcher.
+This data is not meant as a conclusive determination of current Odonate diversity at Woodwalton Fen (52.4433623, -0.2081061), but an incorporation of what people have found at the site in 2019 and 2020.
+Four species were listed on the iNaturalist species list for Woodwalton Fen (https://www.inaturalist.org/places/woodwalton-fen).
+These species were *C.
+puella* (British, 2020), *S.
+sanguineum* (Ueda, 2020), *L.
+quadrimaculata* (Ueda, 2020), and *B.
+pratense* (Ueda, 2020).
+
+This project was inherently limited by the structure and lack of completeness of the data set obtained.
+Perhaps the most important conclusion of this project is the forethought required in how data is collected and databased for future use.
+Morris, the original researcher, collected this information in a time when these considerations did not need to be taken, as open source data sets were certainly not on the average researcher's mind.
+While this study is limited in its analytic ability, the author does not intend to masquerade the results as such.
+Much of the ecological information needed in order to determine the most important factors in *Odonata* diversity (vegetation, pond area, scrub management as possibilities) is not available.
+Therefore, the findings of the original study cannot be expanded upon in an ecological sense.
+It also showcases the difficulty in finding representative and complete data sets, with the admission that this data set might not have been the best choice, at least not on its own.
+
 # Reproducibility
@@
 
-This project was inherently limited by the structure and lack of completeness of the data set obtained.
-Perhaps the most important conclusion of this project is the forethought required in how data is collected and databased for future use.
-Morris, the original researcher, collected this information in a time when these considerations did not need to be taken, as open source data sets were certainly not on the average researcher's mind.
-While this study is limited in its analytical ability, the author does not intend to masquerade the results as such.
-Much of the ecological information needed in order to determine the most important factors in *Odonata* diversity (vegetation, pond area, scrub management) is not available.
-Therefore, the findings of the original study cannot be expanded upon in an ecological sense.
-It also showcases the difficulty in finding representative and complete data sets, with the admission that this data set might not have been the best choice, at least not on its own.
+The construction of the area as a flood reservoir was successful in increasing Odonate species at the fen and in increasing the overall abundance of Odonates in the area.
+Additionally, while not certain, there is possibility that the  trend is that of a rapid and early colonization by Odonates.
+If true, this could be good news for places around the world with degraded wetlands.
+If fauna and flora can recolonize quickly after changes are made to return land to more closely resemble its original form, we have hope for the reclamation of former cropland.
+Drainage of wetlands or clearance of forests for agricultural use is widespread both temporally and geographically, with about 1/3 of global land under agricultural use (DeFries et al.
+2004).
+With many nations and localities considering the environmental trade-offs of land-use for agriculture, success stories like the Odonates in Woodwalton Fen are beneficial.
+
+These fens (as a type of wetland) provide multiple ecosystem services, biodiversity, water quality management, flood management, and carbon storage (Zedler & Kercher, 2005).
 
 # References
+
+British Dragonfly Society Recording Scheme (2020).
+Dragonfly records from the British Dragonfly Society Recording Scheme.
+Occurrence dataset https://doi.org/10.15468/cuyjyi accessed via GBIF.org on 2020-12-03.
 
@@
 
+Morris, J., Gowing D.
+J.
+G., Mills, J., Dunderdale, J.
+A.
+L.
+(2000) Reconciling agricultural economic and environmental objectives: the case of recreating wetlands in the Fenland area of Eastern England.
+*Agriculture, Ecosystems and Environment*, **79**, 245-257.
+
+Ueda K (2020). iNaturalist Research-grade Observations. iNaturalist.org. Occurrence dataset https://doi.org/10.15468/ab3s5x accessed via GBIF.org on 2020-12-03
+
 Waage, J. K. (1979) Dual function of the damselfly penis: Sperm removal and transfer. *Science*, **203**:4383, 916-918.
@@
 The Wildlife Trusts (n.d.) About the Great Fen. https://www.greatfen.org.uk/about-great-fen
+
+Zedler, J. B., Kercher, S. (2005) Wetland Resources: Status, trends, ecosystem services, and restorability. *Annu. Rev. Environ. Resour.*, **30**, 39-74.

### 2020-12-03 manuscript/manuscript.Rmd
more reference and discussion changes

@@
 The abundance corresponds to the sum of the ten species in Table 1 and the p-value for the model was 0.000162.
+Figure 3 is the ten LM of the corresponding species, with p-values shown in Table 2.
+Figure 4 is a model of the total number of species per year over the study period.
+It does not have a fitted line or a p-value for reasons that will be discussed later.
 
@@
 These fens (as a type of wetland) provide multiple ecosystem services, biodiversity, water quality management, flood management, and carbon storage (Zedler & Kercher, 2005).
+They are also excellent resources for wildlife hobbyists and local residents as nature reserves.
+Local residents may be particularly important in the preservation and conservation of fens, as they are the humans most immediately affected by ecosystem changes.
 

### 2020-12-03 manuscript/manuscript.Rmd
README updated with proposal information; tweaks to manuscript
working on Travis CI

@@
 
-```{r, echo=FALSE}
+```{r 1, echo=FALSE}
 install.packages('plyr')
@@
 The first three main objectives will be answered from the dataset from the original study (Morris, 1991).
-The final sub-bullet is not meant to be answered with true rigor, but as an interesting way to include citizen science.
+The final sub-bullet is not meant to be answered with true rigor and was not an original objective proposed, but is an interesting way to include citizen science.
 It is not expected to be comprehensive or a full picture of current Odonate diversity at the study site.
@@
 
-```{r, echo=FALSE}
+```{r 2, echo=FALSE}
 library('plyr')
@@
 
-```{r, echo=FALSE,}
+```{r 3, echo=FALSE,}
 
@@
 
-```{r, echo=TRUE}
+```{r 4, echo=TRUE}
 fitTotal <- lm(abunYearsSums$total_abundance ~ abunYearsSums$year)
@@
 
-```{r, echo=FALSE}
+```{r 5, echo=FALSE}
 library('plyr')
@@
 
-```{r, echo=FALSE, fig.cap="Figure 4.
+```{r 6, echo=FALSE, fig.cap="Figure 4.
 Number of species in each year of the study period.

### 2020-12-03 tests/tests/test-6.R

@@ -1,0 +1,14 @@
+---
+title: "test_1.R"
+output: pdf_document
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
+
+context("6")
+
+test_that("6", 
+	expect_that(length(fullAbundanceDF) = 11)
+)

### 2020-12-03 analysis/analysisSS.Rmd
typo in figure, needed to by fixed in analysis, manuscript and .png replaced with corrected spelling

@@
 
-plot(pNymphulaCompAY$YEAR, pNymphulaCompAY$pNymphulaAbundance, main = "Pyrrhosa nymphula", xlab = "Year", ylab = "Abundance")
+plot(pNymphulaCompAY$YEAR, pNymphulaCompAY$pNymphulaAbundance, main = "Pyrrhosoma nymphula", xlab = "Year", ylab = "Abundance")
 abline(lm (pNymphulaCompAY$pNymphulaAbundance ~ pNymphulaCompAY$YEAR), col="red")
